---
title: "Machine Learning Fundamentals"
author: "Gani Cem Türemen"
---

# Challenge Summary

Your organization wants to know which companies are similar to each other to help in identifying potential customers of a SAAS software solution (e.g. Salesforce CRM or equivalent) in various segments of the market. The Sales Department is very interested in this analysis, which will help them more easily penetrate various market segments.

# Objectives

Apply your knowledge on K-Means and UMAP along with `dplyr`, `ggplot2`, and `purrr` to create a visualization that identifies subgroups in the S&P 500 Index. You will specifically apply:

- Modeling: `kmeans()` and `umap()`
- Iteration: `purrr`
- Data Manipulation: `dplyr`, `tidyr`, and `tibble`
- Visualization: `ggplot2` (bonus `plotly`)

# Solution

## Import Libraries

```{r}
#| output: "false"
library(dplyr)
library(tidyverse)
library(tidyquant)
library(broom)
library(umap)
library(ggrepel)
```

## Read Data

```{r}
sp_500_prices_tbl <- read_rds("~/GitHub/ss23-bdml-gcturemen/code_drafts/data/sp_500_prices_tbl.rds")
sp_500_index_tbl <- read_rds("~/GitHub/ss23-bdml-gcturemen/code_drafts/data/sp_500_index_tbl.rds")
```

## Convert stock prices to a standardized format (daily returns)

```{r}
sp_500_prices_tbl <- sp_500_prices_tbl %>% 
  filter(date >= '2018-01-01') %>%
  group_by(symbol) %>%
  mutate(lag = lag(adjusted)) %>%
  na.omit() %>%
  mutate(diff = adjusted - lag) %>%
  mutate(pct_return = diff / lag)

sp_500_daily_returns_tbl <- sp_500_prices_tbl %>% 
  select(symbol, date, pct_return)

sp_500_daily_returns_tbl %>% 
  glimpse()
```

## Convert to User-Item Format

```{r}
stock_date_matrix_tbl <- sp_500_daily_returns_tbl %>% 
  pivot_wider(names_from = date, values_from = pct_return, values_fill = 0) %>%
  ungroup()

stock_date_matrix_tbl
```

## Perform K-Means Clustering

```{r}
kmeans_obj <- stock_date_matrix_tbl %>%
  select(-symbol) %>%
  kmeans(centers = 4, nstart = 20)

kmeans_obj %>% 
  glance()
```

## Find the optimal value of K
Custom function `kmeans_mapper()`:
```{r}
kmeans_mapper <- function(center = 3) {
    stock_date_matrix_tbl %>%
        select(-symbol) %>%
        kmeans(centers = center, nstart = 20)
}
```
Applying `kmeans_mapper()` and `glance()` functions iteratively:
```{r}
#| warning: false
k_means_mapped_tbl <- tibble(centers = 1:30) %>%
  mutate(k_means = centers %>% map(kmeans_mapper)) %>%
  mutate(glance  = k_means %>% map(glance))
```
Visualizing the result as a __Scree Plot:__
```{r}
k_means_mapped_tbl %>%
  unnest(glance) %>%
  select(centers, tot.withinss) %>%
  
  ggplot(aes(centers, tot.withinss)) +
  geom_point(color = "#2DC6D6", size = 2) +
  geom_line(color = "#2DC6D6", linewidth = 1) +
  ggrepel::geom_label_repel(aes(label = centers), color = "#FF8014", fill = "#282A36", size = 3, label.padding = 0.10, nudge_x = 0.75, nudge_y = 1) + 
  
  labs(title = "Scree Plot",
       subtitle = "Total within-cluster sum of squares for various number of centers",
       caption = "Conclusion: Based on the Scree Plot, a value between 5-10 can be chosen for the number of centers.") +
  theme(
    panel.background = element_rect(fill = "#222222"),
    panel.border = element_rect(color = "white", fill = NA),
    plot.background = element_rect(fill = "#222222"),
    #panel.grid = element_line(color = "white"),
    #legend.background = element_rect(fill='transparent'),
    #legend.box.background = element_rect(fill='transparent')
    line = element_line(color = "white"),
    text = element_text(color = "white"),
    axis.ticks = element_line(color = "white"),
    axis.text = element_text(color = "white")
  )
```

## Apply UMAP

Applying `umap()` function to the `stock_date_matrix_tbl`:
```{r}
umap_results <- stock_date_matrix_tbl %>% 
  select(-symbol) %>% 
  umap()
```
Combining the `layout` from the `umap_results` with the `symbol` column from the `stock_date_matrix_tbl`:
```{r}
#| output: false
umap_results_tbl <- umap_results$layout %>%
  as_tibble(.name_repair = "unique") %>%
  set_names(c("x", "y")) %>%
  bind_cols(stock_date_matrix_tbl %>% select(symbol))
```
Visualizing the `umap_results_tbl`:
```{r}
#| warning: false
umap_results_tbl %>%
  ggplot(aes(x, y)) +
  geom_point(alpha = 0.5, size = 0.5, color = '#2DC6D6') + 
  geom_label_repel(aes(label = symbol), size = 1, max.overlaps = 20, label.padding = 0.10, color = '#FF8014', seed = 777, nudge_x = 0.25, nudge_y = 0.05, fill = "#282A36") + 
  #theme_tq() + 
  labs(title = 'UMAP 2D Projection') +
  theme(
    panel.background = element_rect(fill = "#222222"),
    panel.border = element_rect(color = "white", fill = NA),
    plot.background = element_rect(fill = "#222222"),
    line = element_line(color = "white"),
    text = element_text(color = "white"),
    axis.ticks = element_line(color = "white"),
    axis.text = element_text(color = "white")
  )
```

## Combine K-Means and UMAP

Pulling out the K-Means for 10 centers:
```{r}
k_means_obj <- k_means_mapped_tbl %>%
  pull(k_means) %>%
  pluck(10)
```
Combining the clusters from `k_means_obj` with `umap_results_tbl`:
```{r}
k_means_10_clusters_tbl <- k_means_obj %>% 
  augment(stock_date_matrix_tbl) %>%
  select(symbol, .cluster)

umap_kmeans_results_tbl <-  k_means_10_clusters_tbl %>%
  left_join(umap_results_tbl, by = 'symbol') %>% 
  left_join(sp_500_index_tbl %>% select(symbol, company, sector), by = 'symbol')
  
```
Visualizing the K-Means and UMAP results:
```{r}
#| warning: false
umap_kmeans_results_tbl %>%
  mutate(label_text = str_glue("Symbol: {symbol}
                                 Cluster: {.cluster}")) %>%
  
  ggplot(aes(x, y, color = .cluster)) +
  
  # Geometries
  geom_point(alpha = 0.5, size = 0.5) +
  geom_label_repel(aes(label = label_text), size = 1,max.overlaps = 20, fill = "#282A36", label.padding = 0.10, nudge_x = 0.25, nudge_y = 0.05) +
  
  # Formatting
  scale_color_discrete() +
  labs(title = "Company Segmentation: 2D Projection",
       subtitle = "UMAP 2D Projection with K-Means Cluster Assignment") +
  theme(
    legend.position = "none",
    #legend.key = element_rect(fill = umap_kmeans_results_tbl$.cluster),
    panel.background = element_rect(fill = "#222222"),
    panel.border = element_rect(color = "white", fill = NA),
    legend.background = element_rect(fill = "#222222"),
    plot.background = element_rect(fill = "#222222"),
    line = element_line(color = "white"),
    text = element_text(color = "white"),
    axis.ticks = element_line(color = "white"),
    axis.text = element_text(color = "white")
  )
```