---
title: "Automated Machine Learning with H2O (I)"
author: "Gani Cem Türemen"
---

# Challenge Summary

Attrition is a problem that impacts all businesses, irrespective of geography, industry and size of the company. Employee attrition leads to significant costs for a business, including the cost of business disruption, hiring new staff and training new staff.

We will use classification models to predict if an employee is likely to quit so that the situation can be intervened on time and remedied to prevent attrition.

# Objectives

We will use our learning from descriptive features and `plot_ggpairs()` to further investigate the features of the employee attrition dataset and answer the related questions.

# Solution

## Import Libraries

```{r}
#| warning: false

library(tidyverse)
library(GGally)
```

## Read Data

```{r}
#| output: false

employee_attrition_tbl <- read_csv(
  "data/datasets-1067-1925-WA_Fn-UseC_-HR-Employee-Attrition.csv")
```

```{r}
#| echo: false

employee_attrition_tbl %>% glimpse()
```

```{r}
#| include: false
# Plot helper function

plot_ggpairs <- function(data, color = NULL, density_alpha = 0.5) {
  
  color_expr <- enquo(color)
  
  if (rlang::quo_is_null(color_expr)) {
    g <- data %>% ggpairs(lower = "blank") 
  } else {
    color_name <- quo_name(color_expr)
    g <- data %>% ggpairs(mapping = aes_string(color = color_name), 
                          lower = "blank", 
                          legend = 1,
                          diag = list(continuous = wrap("densityDiag", 
                                                        alpha = density_alpha))) + 
      theme(legend.position = "bottom", 
            legend.background = element_rect(fill = NA, color = 'white'),
            legend.key = element_blank(),
            plot.background = element_rect(fill = "#222222", color = "#222222"),
            text = element_text(color = "white"),
            axis.ticks = element_line(color = "white"),
            axis.text = element_text(color = "white"))
  }
  return(g)
}
```

## Compensation Features

```{r}
#| echo: false
#| warning: false

employee_attrition_tbl %>% select(Attrition, 
                                  contains("income"), 
                                  contains("rate"), 
                                  contains("salary"), 
                                  contains("stock")) %>% 
  plot_ggpairs(Attrition)
```

__Question:__ What can you deduce about the interaction between Monthly Income and Attrition?

    a. Those that are leaving the company have a higher Monthly Income
    b. That those are staying have a lower Monthly Income
    c. Those that are leaving have a lower Monthly Income
    d. It's difficult to deduce anything based on the visualization

__Answer:__ Looking at the second figure in the diagonal it can be seen that the employees that are leaving have a peak at a lower monthly income value compared to those that are staying. The answer is c.

__Question:__ What can you deduce about the interaction between Percent Salary Hike and Attrition?

    a. Those that are leaving the company have a higher Percent Salary Hike
    b. Those that are staying have a lower Percent Salary Hike
    c. Those that are leaving have lower Percent Salary Hike
    d. It's difficult to deduce anything based on the visualization

__Answer:__ Looking at the sixth figure in the diagonal, it is seen that distributions for both employee groups are similar so it is difficult to make any comments. The answer is d.

__Question:__ What can you deduce about the interaction between Stock Option Level and Attrition?

    a. Those that are leaving the company have a higher stock option level
    b. Those that are staying have a higher stock option level
    c. It's difficult to deduce anything based on the visualization

__Answer:__ Since the number of staying employees are concentrated around higher stock option levels (1, 2) compared to the employees that are leaving (0) the answer is b.

## Survey results

```{r}
#| echo: false
#| warning: false

employee_attrition_tbl %>% select(Attrition, 
                                  contains("satisfaction"), 
                                  contains("life")) %>% 
  plot_ggpairs(Attrition)
```

__Question:__ What can you deduce about the interaction between Environment Satisfaction and Attrition?

    a. A higher proportion of those leaving have a low environment satisfaction level
    b. A higher proportion of those leaving have a high environment satisfaction level
    c. It's difficult to deduce anything based on the visualization
    
__Answer:__ It is difficult to make a comment since it is not clearly defined what 'high' and 'low' satisfaction are. The answer is c.

__Question:__ What can you deduce about the interaction between Work Life Balance and Attrition

    a. Those that are leaving have higher density of 2's and 3's
    b. Those that are staying have a higher density of 2's and 3's
    c. Those that are staying have a lower density of 2's and 3's
    d. It's difficult to deduce anything based on the visualization
    
__Answer:__ From the fifth figure in the diagonal, it can be seen that staying employees have a higher density of 2's and 3's. The answer is b.

## Performance Data

```{r}
#| echo: false
#| warning: false

employee_attrition_tbl %>% select(Attrition, 
                                  contains("performance"), 
                                  contains("involvement")) %>% 
  plot_ggpairs(Attrition)
```

__Question:__ What can you deduce about the interaction between Job Involvement and Attrition?

    a. Those that are leaving have a lower density of 3's and 4's
    b. Those that are leaving have a lower density of 1's and 2's
    c. Those that are staying have a lower density of 2's and 3's
    d. It's difficult to deduce anything based on the visualization
    
__Answer:__ From the figure, it can be seen that leaving employees have a lower density of 3's and 4's. The answer is a.

## Work-Life Features

```{r}
#| echo: false
#| warning: false

employee_attrition_tbl %>% select(Attrition, 
                                  contains("overtime"), 
                                  contains("travel")) %>% 
  plot_ggpairs(Attrition)
```

__Question:__ What can you deduce about the interaction between Over Time and Attrition?

    a. The proportion of those leaving that are working Over Time are high compared to those that are not leaving
    b. The proportion of those staying that are working Over Time are high compared to those that are not staying
    
__Answer:__ Regarding Over Time, proportion of staying employees is higher in both groups. The answer is b.

## Training and Education

```{r}
#| echo: false
#| warning: false

employee_attrition_tbl %>% select(Attrition, 
                                  contains("training"), 
                                  contains("education")) %>% 
  plot_ggpairs(Attrition)
```

__Question:__ What can you deduce about the interaction between Training Times Last Year and Attrition

    a. People that leave tend to have more annual trainings
    b. People that leave tend to have less annual trainings
    c. It's difficult to deduce anything based on the visualization
    
__Answer:__ From the figure it is seen that distributions of both groups concentrate around similar values. The answer is c.

## Time-Based Features

```{r}
#| echo: false
#| warning: false

employee_attrition_tbl %>% select(Attrition, contains("years")) %>% 
  plot_ggpairs(Attrition)
```

__Question:__ What can you deduce about the interaction between Years At Company and Attrition

    a. People that leave tend to have more working years at the company
    b. People that leave tend to have less working years at the company
    c. It's difficult to deduce anything based on the visualization

__Answer:__ It is apparent that people that leave tend to have less working years at the company because corresponding distribution peaks at a smaller value. The answer is b.

__Question:__ What can you deduce about the interaction between Years Since Last Promotion and Attrition?

    a. Those that are leaving have more years since last promotion than those that are staying
    b. Those that are leaving have fewer years since last promotion than those that are staying
    c. It's difficult to deduce anything based on the visualization
    
__Answer:__ Since both distributions peak at the lower end of the x axis and they are similar in other parts it is hard to make a decision. The answer is c.