---
title: "Automated Machine Learning with H2O (II)"
author: "Gani Cem Türemen"
---

# Challenge Summary

For the challenge, we shall be working with a Product Backorders dataset. The goal here is to predict whether or not a product will be put on backorder status, given a number of product metrics such as current inventory, transit time, demand forecasts and prior sales. It’s a classic Binary Classification problem.

# Objectives

We will apply our learning of H2O Automated Machine Learning to predict backorder status.

# Solution

## Import Libraries

```{r}
#| output: false

library(tidyverse)
library(recipes)
library(rsample)
library(readxl)
library(h2o)
```

## Load the training & test dataset

```{r}
#| output: false

product_backorders_tbl <- read_csv("data/product_backorders.csv")

product_backorders_tbl %>% glimpse()
```

```{r}
#| echo: false

product_backorders_tbl %>% glimpse()
```

Split the data into training and test sets:
```{r}
set.seed(seed = 777)
split_obj <- initial_split(product_backorders_tbl, prop = 0.80)

train_set_tbl <- training(split_obj)
test_set_tbl  <- testing(split_obj)
```

Since some columns such as `deck_risk` are not numeric, they have to be converted into factor. The recipe below will take care of that:
```{r}
recipe_obj <- recipe(went_on_backorder ~., data = train_set_tbl) %>% 
  step_zv(all_predictors()) %>% 
  step_mutate_at(potential_issue,
                 deck_risk,
                 oe_constraint,
                 ppap_risk,
                 stop_auto_buy,
                 rev_stop,
                 went_on_backorder,
                 fn = as.factor) %>% 
  prep()

train_tbl <- bake(recipe_obj, new_data = train_set_tbl)
test_tbl  <- bake(recipe_obj, new_data = test_set_tbl)

train_tbl %>% glimpse()
```

## Specify the response and predictor variables

To determine the backorder status `went_on_backorder`, all the other columns will be used as predictors:
```{r}
#| output: false

h2o.init()

split_h2o <- h2o.splitFrame(as.h2o(train_tbl), ratios = c(0.80), seed = 767)
train_h2o <- split_h2o[[1]]
valid_h2o <- split_h2o[[2]]
test_h2o  <- as.h2o(test_tbl)

y <- "went_on_backorder"
x <- setdiff(names(train_h2o), y)
```
    
## Run AutoML specifying the stopping criterion

Set 30 seconds for `max_runtime_secs` and 5 for `nfolds`:
```{r}
#| output: false

automl_models_h2o <- h2o.automl(
  x = x,
  y = y,
  training_frame    = train_h2o,
  validation_frame  = valid_h2o,
  leaderboard_frame = test_h2o,
  max_runtime_secs  = 30,
  nfolds            = 5 
)
```

## View the leaderboard

```{r}
automl_models_h2o@leaderboard
```

## Predicting using Leader Model

```{r}
#| include: false
# Extracts and H2O model name by a position so can more easily use h2o.getModel()

extract_h2o_model_name_by_position <- function(h2o_leaderboard, n = 1, verbose = T) { 
  
  model_name <- h2o_leaderboard %>% as_tibble() %>% 
    slice(n) %>% 
    pull(model_id)
    
    if (verbose) message(model_name)
    
    return(model_name)
    
}
```

Extract the first model of the leaderboard:
```{r}
leader_model <- automl_models_h2o@leaderboard %>% 
  extract_h2o_model_name_by_position(1) %>% 
  h2o.getModel()
```

Use the extracted model to make predictions on the test set:
```{r}
#| output: false

predictions <- h2o.predict(leader_model, newdata = as.h2o(test_tbl))

predictions_tbl <- predictions %>% as_tibble()

predictions_tbl %>% glimpse()
```

```{r}
#| echo: false

predictions_tbl %>% glimpse()
```

## Save the leader model

Save the extracted model in case it is needed later:
```{r}
#| output: false

leader_model %>% h2o.saveModel(path = "h2o_models/")
```

```{r}
#| include: false

h2o.shutdown()
```