---
title: "Performance Measures"
author: "Gani Cem Türemen"
---

# Challenge Summary

For the challenge, we will continue working on the Product Backorders dataset. The goal here is to visualize various performance metrics of H2O models to simplify evaluation of them.

# Objectives

We will apply our learning of H2O Automated Machine Learning to predict backorder status and visualize performance metrics to evaluate models.

# Solution

## Import Libraries

```{r}
#| output: false

library(tidyverse)
library(readxl)
library(rsample)
library(h2o)
library(recipes)
library(cowplot)
library(glue)
```

## Load the training & test dataset

```{r}
#| output: false

product_backorders_tbl <- read_csv("data/product_backorders.csv")

product_backorders_tbl %>% glimpse()
```

```{r}
#| echo: false

product_backorders_tbl %>% glimpse()
```

Split the data into training and test sets:
```{r}
set.seed(seed = 777)
split_obj <- initial_split(product_backorders_tbl, prop = 0.80)

train_set_tbl <- training(split_obj)
test_set_tbl  <- testing(split_obj)
```

Since some columns such as `deck_risk` are not numeric, they have to be converted into factor. The recipe below will take care of that:
```{r}
recipe_obj <- recipe(went_on_backorder ~., data = train_set_tbl) %>% 
  step_zv(all_predictors()) %>% 
  step_mutate_at(potential_issue,
                 deck_risk,
                 oe_constraint,
                 ppap_risk,
                 stop_auto_buy,
                 rev_stop,
                 went_on_backorder,
                 fn = as.factor) %>% 
  prep()

train_tbl <- bake(recipe_obj, new_data = train_set_tbl)
test_tbl  <- bake(recipe_obj, new_data = test_set_tbl)

train_tbl %>% glimpse()
```

## Specify the response and predictor variables

To determine the backorder status `went_on_backorder`, all the other columns will be used as predictors:
```{r}
#| output: false

h2o.init()

split_h2o <- h2o.splitFrame(as.h2o(train_tbl), ratios = c(0.80), seed = 767)
train_h2o <- split_h2o[[1]]
valid_h2o <- split_h2o[[2]]
test_h2o  <- as.h2o(test_tbl)

y <- "went_on_backorder"
x <- setdiff(names(train_h2o), y)
```
    
## Run AutoML

Set 600 seconds for `max_runtime_secs` and 15 for `max_models` to limit the computation time:
```{r}
#| output: false

automl_models_h2o <- h2o.automl(
  x = x,
  y = y,
  training_frame    = train_h2o,
  validation_frame  = valid_h2o,
  leaderboard_frame = test_h2o,
  max_runtime_secs  = 600,
  max_models = 15,
  nfolds            = 5,
  include_algos = c("DeepLearning", "GBM", "XGBoost", "GLM", "DRF")
)
```

## View the leaderboard

```{r}
automl_models_h2o@leaderboard
```

## Leaderboard visualization

First, leaderboard will be converted into long format to be able to visualize it easier:
```{r}
model_leaderboard <- automl_models_h2o@leaderboard %>%
  as_tibble() %>%
  select(-c(aucpr, mean_per_class_error, rmse, mse)) %>%
  mutate(model_type = str_extract(model_id, "[^_]+")) %>%
  rownames_to_column(var = "row_name") %>%
  mutate(model_id   = as_factor(model_id) %>% reorder(auc), 
         model_type = as.factor(model_type)) %>% 
  pivot_longer(cols = -c(model_id, model_type, row_name), 
               names_to = "key", 
               values_to = "value", 
               names_transform = list(key = forcats::fct_inorder)) %>% 
  mutate(model_id = paste0(row_name, ". ", model_id) %>% as_factor() %>% fct_rev())

model_leaderboard %>% glimpse()
```

Visualize the long format leaderboard:
```{r}
model_leaderboard %>%
  ggplot(aes(value, model_id, color = model_type)) +
  geom_point(size = 3) +
  geom_label(aes(label = round(value, 3), hjust = "inward"), label.padding = unit(0.10, "lines"), size = 3) +
  facet_wrap(~ key, scales = "free_x") +
  labs(title = "Leaderboard Metrics",
       subtitle = paste0("Ordered by: ", "auc"),
       y = "Model Position, Model ID",
       x = "",
       col = "Model Type") + 
  theme(
    legend.position = "bottom",
    legend.background = element_rect(fill = NA, color = "white"),
    panel.background = element_blank(),
    panel.border = element_rect(color = "white", fill = NA),
    plot.background = element_rect(fill = "#222222", color = "#222222"),
    line = element_line(color = "white"),
    text = element_text(color = "white", size = 7),
    axis.ticks = element_line(color = "white"),
    axis.text = element_text(color = "white", size = 6),
  )
```

From the figure, it is seen that `AUC` and `LogLoss` metrics behave mostly as expected. Generally with increasing model performance, an increase in `AUC` and a decrease in `LogLoss` is observed.

## Tune a model with grid search

```{r}
#| include: false
# Extracts and H2O model name by a position so can more easily use h2o.getModel()

extract_h2o_model_name_by_position <- function(h2o_leaderboard, n = 1, verbose = T) {
    
    model_name <- h2o_leaderboard %>%
        as_tibble() %>%
        slice(n) %>%
        pull(model_id)
    
    if (verbose) message(model_name)
    
    return(model_name)
    
}
```

A `Deep Learning` model will be tuned in this section. First, the Deep Learning model from the leaderboard is extracted to be used for comparison purposes:
```{r}
dl_model_automl <- automl_models_h2o@leaderboard %>% 
  extract_h2o_model_name_by_position(11) %>% 
  h2o.getModel()
```

```{r}
#| output: false

p_dl_model_automl <- h2o.performance(dl_model_automl, newdata = as.h2o(test_tbl))

p_dl_model_automl
```

```{r}
#| echo: false

p_dl_model_automl
```

```{r}
#| include: false

p_dl_model_automl_tbl <- p_dl_model_automl %>% 
  h2o.metric() %>%
  as_tibble() %>%
  mutate(auc = h2o.auc(p_dl_model_automl))
```

It is seen that this model has an `AUC` value of 0.775 and a `LogLoss` value of 0.321. Using grid search, there is a possibility to find a model with better performance metrics.

Use the hyperparameters `hidden` and `epochs` for tuning:

    hidden: Hidden layer size. 
    lambda: The number of times to iterate the data set.
    
```{r}
#| warning: false
#| output : false

model_grid <- h2o.grid(
  algorithm = "deeplearning",
  grid_id = "deeplearning_grid_01",
  x = x,
  y = y,
  training_frame = train_h2o,
  validation_frame = valid_h2o,
  max_runtime_secs  = 600,
  nfolds = 5,
  hyper_params = list(
    #activation = c("tanh", "rectifier"),
    hidden = list(c(10, 10), c(15, 15), c(10, 10, 10), c(15, 15, 15)),
    epochs = c(50, 100, 200)
  )
)
```

Get the grid search result and order models by decreasing AUC: 
```{r}
#| warning: false

h2o.getGrid(grid_id = "deeplearning_grid_01", sort_by = "auc", decreasing = TRUE)
```

Extract the leader model of the grid search: 
```{r}
dl_model_grid_search <- h2o.getModel("deeplearning_grid_01_model_9")
```

```{r}
dl_model_grid_search %>% h2o.auc(train = T, valid = T, xval = T)
```

No overfitting is observed since the AUC value for training, validation and cross-validation are all similar.

```{r}
#| output: false

p_dl_model_grid_search <- dl_model_grid_search %>% h2o.performance(newdata = as.h2o(test_tbl))

p_dl_model_grid_search
```

```{r}
#| echo: false

p_dl_model_grid_search
```

It is seen that by using grid search it was possible find a model with improved performance metrics compared to the base model found in the previous section.

|             Model               |  AUC  |  LogLoss  |
|:-------------------------------:|:-----:|:---------:|
| dl_model_automl                 | 0.775 |   0.321   |
| dl_model_grid_search            | 0.895 |   0.263   |

## Visualize the trade-off between the precision and the recall and the optimal threshold

Convert the performance object of the model into a series of different metrics which vary by threshold and store it as `p_dl_model_grid_search_tbl`:
```{r}
p_dl_model_grid_search_tbl <- p_dl_model_grid_search %>% 
  h2o.metric() %>%
  as_tibble() %>%
  mutate(auc = h2o.auc(p_dl_model_grid_search))
```

Visualize `precision` and `recall` by selecting corresponding columns:
```{r}
#| warning: false
#| output: false

p_dl_model_grid_search_tbl %>% select(c(threshold, precision, recall)) %>% 
  pivot_longer(cols = c(precision, recall), names_to = "key", values_to = "values") %>%
  
  ggplot(aes(x = threshold, y = values, color = key)) +
  geom_line(size = 1) +
  scale_color_manual(values = c("#2DC6D6", "#FF8014")) +
  geom_vline(xintercept = h2o.find_threshold_by_max_metric(p_dl_model_grid_search, "f1"), 
             color = 'red', size = 1) +
  labs(title = "Precision vs Recall", subtitle = "dl_model_grid_search", y = "Value", x = "Threshold", col = "") +
  theme(
    legend.position = "bottom",
    legend.background = element_rect(fill = NA, color = "white"),
    legend.key = element_blank(),
    panel.background = element_blank(),
    panel.border = element_rect(color = "white", fill = NA),
    plot.background = element_rect(fill = "#222222", color = "#222222"),
    line = element_line(color = "white"),
    text = element_text(color = "white"),
    axis.ticks = element_line(color = "white"),
    axis.text = element_text(color = "white")
  )
```

```{r}
#| echo: false
#| warning: false

p_dl_model_grid_search_tbl %>% select(c(threshold, precision, recall)) %>% 
  pivot_longer(cols = c(precision, recall), names_to = "key", values_to = "values") %>%
  
  ggplot(aes(x = threshold, y = values, color = key)) +
  geom_line(size = 1) +
  scale_color_manual(values = c("#2DC6D6", "#FF8014")) +
  geom_vline(xintercept = h2o.find_threshold_by_max_metric(p_dl_model_grid_search, "f1"), 
             color = 'red', size = 1) +
  labs(title = "Precision vs Recall", subtitle = "dl_model_grid_search", y = "Value", x = "Threshold", col = "") +
  theme(
    legend.position = "bottom",
    legend.background = element_rect(fill = NA, color = "white"),
    legend.key = element_blank(),
    panel.background = element_blank(),
    panel.border = element_rect(color = "white", fill = NA),
    plot.background = element_rect(fill = "#222222", color = "#222222"),
    line = element_line(color = "white"),
    text = element_text(color = "white"),
    axis.ticks = element_line(color = "white"),
    axis.text = element_text(color = "white")
  )
```

From the figure, values of `precision` and `recall` can be seen as the threshold value increases. The vertical red line indicates the optimal threshold value which maximizes the `F1`.

## ROC Plot

```{r}
#| include: false

p_dl_model_automl_tbl <- p_dl_model_automl_tbl %>% 
  mutate(auc = auc %>% round(3) %>% as.character() %>% as_factor())

p_dl_model_grid_search_tbl <- p_dl_model_grid_search_tbl %>% 
  mutate(auc  = auc %>% round(3) %>% as.character() %>% as_factor())
```

```{r}
roc_plot <- ggplot() + 
  geom_line(data = p_dl_model_automl_tbl, 
            aes(x = fpr, y = tpr, color = "dl_model_automl"), 
            size = 1) +
  geom_line(data = p_dl_model_grid_search_tbl, 
            aes(x = fpr, y = tpr, color = "dl_model_grid_search"), 
            size = 1) +
  geom_abline(color = "red", 
              linetype = "dotted", 
              size = 1) +
  scale_color_manual(name = "", 
                     values = c("dl_model_automl" = "#2DC6D6", "dl_model_grid_search" = "#FF8014")) +
  labs(title = "ROC Plot", 
       subtitle = "Performance Comparison of dl_model_automl and dl_model_grid_search", 
       x = "False Positive Rate (FPR)", 
       y = "True Positive Rate (TPR)") +
  theme(
    legend.position = "bottom",
    legend.background = element_rect(fill = NA, color = "white"),
    legend.key = element_blank(),
    panel.background = element_blank(),
    panel.border = element_rect(color = "white", fill = NA),
    plot.background = element_rect(fill = "#222222", color = "#222222"),
    line = element_line(color = "white"),
    text = element_text(color = "white"),
    axis.ticks = element_line(color = "white"),
    axis.text = element_text(color = "white")
  )

roc_plot
```

It is also seen from the ROC plot that the model found with grid search (`dl_model_grid_search`) has better performance since it has a larger area under its curve compared to the model that has been found automatically (`dl_model_automl`). 

## Precision vs Recall Plot

```{r}
pr_plot <- ggplot() +
  geom_line(data = p_dl_model_automl_tbl, 
            aes(x = recall, y = precision, color = "dl_model_automl"), 
            size = 1) +
  geom_line(data = p_dl_model_grid_search_tbl, 
            aes(x = recall, y = precision, color = "dl_model_grid_search"), 
            size = 1) +
  scale_color_manual(name = "", 
                     values = c("dl_model_automl" = "#2DC6D6", "dl_model_grid_search" = "#FF8014")) +
  labs(title = "Precision vs Recall Plot", 
       subtitle = "Performance Comparison of dl_model_automl and dl_model_grid_search", 
       x = "Recall", 
       y = "Precision") +
  theme(
    legend.position = "bottom",
    legend.background = element_rect(fill = NA, color = "white"),
    legend.key = element_blank(),
    panel.background = element_blank(),
    panel.border = element_rect(color = "white", fill = NA),
    plot.background = element_rect(fill = "#222222", color = "#222222"),
    line = element_line(color = "white"),
    text = element_text(color = "white"),
    axis.ticks = element_line(color = "white"),
    axis.text = element_text(color = "white")
  )

pr_plot
```

A similar observation can also be made for the precision-recall curve. It is seen that `dl_model_grid_search` performs better compared to `dl_model_automl` since its curve is closer to the upper right corner of the plot (which means the model can maintain higher precision while recall increases).

## Gain Plot

```{r}
gain_lift_automl_tbl <- p_dl_model_automl %>%
  h2o.gainsLift() %>%
  as_tibble()

gain_automl_transformed_tbl <- gain_lift_automl_tbl %>% 
  select(group, cumulative_data_fraction, cumulative_capture_rate, cumulative_lift) %>%
  select(-contains("lift")) %>%
  mutate(baseline = cumulative_data_fraction) %>%
  rename(gain     = cumulative_capture_rate) %>%
  pivot_longer(cols = c(gain, baseline), values_to = "value", names_to = "key")

gain_lift_grid_tbl <- p_dl_model_grid_search %>%
  h2o.gainsLift() %>%
  as_tibble()

gain_grid_transformed_tbl <- gain_lift_grid_tbl %>% 
  select(group, cumulative_data_fraction, cumulative_capture_rate, cumulative_lift) %>%
  select(-contains("lift")) %>%
  mutate(baseline = cumulative_data_fraction) %>%
  rename(gain     = cumulative_capture_rate) %>%
  pivot_longer(cols = c(gain, baseline), values_to = "value", names_to = "key")
```

```{r}
gain_plot <- ggplot() +
  geom_line(data = gain_automl_transformed_tbl %>% filter(str_detect(key, "gain")), 
            aes(x = cumulative_data_fraction, y = value, color = "dl_model_automl"),
            size = 1) +
  geom_line(data = gain_grid_transformed_tbl %>% filter(str_detect(key, "gain")), 
            aes(x = cumulative_data_fraction, y = value, color = "dl_model_grid_search"),
            size = 1) +
  geom_abline(color = "red", 
            linetype = "dotted", 
            size = 1) +
  scale_color_manual(name = "", 
                     values = c("dl_model_automl" = "#2DC6D6", 
                                "dl_model_grid_search" = "#FF8014")) +
  labs(title = "Gain Chart", x = "Cumulative Data Fraction", y = "Gain") +
  theme(
    legend.position = "bottom",
    legend.background = element_rect(fill = NA, color = "white"),
    legend.key = element_blank(),
    panel.background = element_blank(),
    panel.border = element_rect(color = "white", fill = NA),
    plot.background = element_rect(fill = "#222222", color = "#222222"),
    line = element_line(color = "white"),
    text = element_text(color = "white"),
    axis.ticks = element_line(color = "white"),
    axis.text = element_text(color = "white")
  )

gain_plot
```

From the figure, it can be seen that with `dl_model_grid_search`, the ability to target backorders from a given size of data has increased significantly, especially at the lower end of the x-axis (where it is important).

## Lift Plot

```{r}
lift_automl_transformed_tbl <- gain_lift_automl_tbl %>%
  select(group, cumulative_data_fraction, cumulative_capture_rate, cumulative_lift) %>%
  select(-contains("capture")) %>%
  mutate(baseline = 1) %>%
  rename(lift = cumulative_lift) %>%
  pivot_longer(cols = c(lift, baseline), values_to = "value", names_to = "key")

lift_grid_transformed_tbl <- gain_lift_grid_tbl %>% 
  select(group, cumulative_data_fraction, cumulative_capture_rate, cumulative_lift) %>%
  select(-contains("capture")) %>%
  mutate(baseline = 1) %>%
  rename(lift = cumulative_lift) %>%
  pivot_longer(cols = c(lift, baseline), values_to = "value", names_to = "key")
```

```{r}
lift_plot <- ggplot() +
  geom_line(data = lift_automl_transformed_tbl %>% filter(str_detect(key, "lift")), 
            aes(x = cumulative_data_fraction, y = value, color = "dl_model_automl"),
            size = 1) +
  geom_line(data = lift_grid_transformed_tbl %>% filter(str_detect(key, "lift")), 
            aes(x = cumulative_data_fraction, y = value, color = "dl_model_grid_search"),
            size = 1) +
  geom_hline(yintercept = 1,
             color = "red", 
            linetype = "dotted", 
            size = 1) +
  scale_color_manual(name = "", 
                     values = c("dl_model_automl" = "#2DC6D6", 
                                "dl_model_grid_search" = "#FF8014")) +
  labs(title = "Lift Chart", x = "Cumulative Data Fraction", y = "Lift") +
  theme(
    legend.position = "bottom",
    legend.background = element_rect(fill = NA, color = "white"),
    legend.key = element_blank(),
    panel.background = element_blank(),
    panel.border = element_rect(color = "white", fill = NA),
    plot.background = element_rect(fill = "#222222", color = "#222222"),
    line = element_line(color = "white"),
    text = element_text(color = "white"),
    axis.ticks = element_line(color = "white"),
    axis.text = element_text(color = "white")
  )

lift_plot
```

Similarly, it is seen from the figure that `dl_model_grid_search` is better at targeting backorders for the same size of data when compared with `dl_model_automl`.

## Dashboard with cowplot

```{r}
p_legend <- get_legend(roc_plot)

roc_plot <- roc_plot + theme(legend.position = "none") + labs(subtitle = element_blank())
pr_plot <- pr_plot + theme(legend.position = "none") + labs(subtitle = element_blank())
gain_plot <- gain_plot + theme(legend.position = "none")
lift_plot <- lift_plot + theme(legend.position = "none")

p <- plot_grid(roc_plot, pr_plot, gain_plot, lift_plot, ncol = 2)

p_title <- ggdraw() + 
  draw_label("H2O Model Metrics", size = 18, fontface = "bold", color = "white")
  
p_subtitle <- ggdraw() + draw_label(glue(""), size = 10, color = "white")
  
combined_plot <- plot_grid(p_title, 
                           p_subtitle, 
                           p, 
                           p_legend, 
                           ncol = 1, 
                           rel_heights = c(0.05, 0.05, 1, 0.05 * 4)) +
  theme(
    legend.position = "bottom",
    legend.background = element_rect(fill = NA, color = "white"),
    panel.background = element_blank(),
    panel.border = element_rect(color = "#222222", fill = NA),
    plot.background = element_rect(fill = "#222222", color = "#222222"),
    line = element_line(color = "white"),
    text = element_text(color = "white"),
    axis.ticks = element_line(color = "white"),
    axis.text = element_text(color = "white")
  )

combined_plot
```

```{r}
#| include: false

h2o.shutdown()
```